Viscous Transport in Eroding Porous Media

### 2019-10-12 docs/transport/erosion_transport.tex
done up to section 6
up to section 6.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results of dense packings of bodies eroding in
Stokes flow. Moreover, we analyze transport through the eroded
geometries by examining coarse-grained variables.  Each eroding body is
initially a circle with center $\cc_i$ and radius $r_i$.  The center and
radius are chosen at random, and the body is accepted if it is contained
in $[-1,1] \times [-1,1]$ and is sufficiently separated from all other
bodies.  Since the error of our quadrature method is bounded independent
of the geometry, we consider bodies that are separated by less than 10\%
of an arclength spacing.  The randomized method is repeated until the
initial geometry reaches a desired initial porosity.

Once the geometry is initialized, each body has length $L_i = 2\upi
r_i$, and they are each discretized with $N_\iin$ points.  A no-slip
boundary condition is imposed on each eroding body $\gamma_i$.  For all
simulations, we discretize the outer wall $\Gamma$ with $N_\out=1024$
points.  For all but the first example, the Dirichlet boundary condition
on $\Gamma$ is a Hagen-Poiseuille flow, and the flow rate is adjusted at
each time step to maintain a constant pressure drop. Since the fluid
equations are linear, this is achieved by computing the pressure near
the inlet and outlet at each time step, and then scaling the flow rate
appropriately~\citep{qua-moo2018}.

The erosion rate loses regularity at stagnation points, which inevitably
leads to corner formation on the bodies.  As described in
section~\ref{sec:time} and our previous work~\citep{qua-moo2018}, we
ameliorate corner formation by introducing a curvature penalization term
with parameter $\epsilon$ and a Gaussian smoothing step with parameter
$\sigma$.  For all examples, we use the smoothing parameters
$\eps=15/256$ and $\sigma=10/256$, and the time step size is $\Delta t =
10^{-4}$.

In each of the examples, we plot the vorticity in the fluid bulk.  Since
the vorticity and the shear stress agree on surface of the eroding
bodies, the vorticity is useful to visualize the erosion rate.  In
particular, the rate of erosion is largest in regions where the
magnitude of the vorticity is largest.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Consider a single eroding body close to a solid wall with the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. \citet{mit-spa2017} performed a similar
three-dimensional experiment using a second-order quadrature method.
Their initial body is a sphere with its center located $1.5$ radii above
the solid wall.  In two dimensions, the trapezoid rule is spectrally
accurate, and an analogous initial configuration is possible. However,
unless the number of discretization points is increased, the trapezoid
rule becomes obsolete if the eroding body is initialized too close to
the solid wall. Therefore, to simulate erosion near a solid wall at
reasonable levels of discretization, we use our hybrid method that
combines the trapezoid rule and the Barycentric quadrature rule.

We consider an eroding body that is initially a circle of radius
$r=0.4$.  We discretize the eroding body with $N_\iin=256$
discretization points, and its center is chosen so that the distance
between the body and the solid wall is $h$, $h/2$, and $h/10$, where $h
= 2\upi r/N_\iin$.  If we used the trapezoid rule and required an error
that is comparable to the Barycentric quadrature rule, the body with an
initial distance of $h/10$ from the solid wall would require $6,400$
discretization points, and the outer wall would require $50,000$
discretization points.

In figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For all three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop, but
interesting shapes form nonetheless.  For instance, in the top half of
the body, there is a sudden increase in the shear stress, and this leads
to a region of high curvature.  This behavior is also present in three
dimensions~\citep[see figure 7(c)]{mit-spa2017}.  The biggest difference
between the two- and three-dimensional results is the presence of a
recirculation zone.  In three dimensions, there is no recirculation
between the solid wall and the spherical body~\citep{cha-feu2003}, but
recirculation is possible in two dimensions~\citep{chw-wu1975, hig1985}.
To visualize the flow, we plot the vorticity of the final time step from
figure~\ref{fig:NearWall} in figure~\ref{fig:NearWall_vort}.  In these
examples, a small recirculation zone, both in size and magnitude, is
present in the region where the vorticity is smallest

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the average pressure drop from
$x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies
have vanished.  Snapshots of the bodies at four equispaced times are
shown in figure~\ref{fig:Eroding20vort}.  The color is the vorticity,
whose absolute value is proportional to the erosion rate on solid
boundaries.  Initially, several of the eroding bodies are closer to the
outer wall than the $5h$ threshold required for the trapezoid rule to
achieve machine precision.  In particular, the distance between bodies
1, 6, 13, and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and
$1.3h$, respectively, where $h$ is the arclength spacing of the outer
wall $\Gamma$.  In addition, the distance between several pairs of
eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too
small to be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies is accurately resolved to the desired accuracy,
and erosion can be simulated until all the bodies have vanished.

\begin{figure}
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin=256$ points, in a Hagen-Poiseuille
flow.  The four snapshots are evenly spaced in time, and the color is
the fluid vorticity. In the fourth frame, bodies 4, 12, and 17 have
vanished completely, and body 19 has almost completely eroded.}
\end{center}
\end{figure}

Erosion causes the space between nearly touching bodies to quickly
expand, and flat faces develop along the regions of near contact.  This
qualitative behavior is seen in figure~\ref{fig:Eroding20vort} between
bodies 3 \& 4, 15 \& 16, and was also observed in previous
work~\citep{qua-moo2018}. However, by resolving the interaction between
bodies that are much closer together than in previous work, we observe
that very little erosion occurs between certain pairs of bodies, at
least initially.  For instance the opening between bodies 1 \& 6, 3 \&
9, and 5 \& 13 grow much slower than the opening between bodies 15 \&
16.  A common feature of the pores that grow slowly is that they are
nearly perpendicular to the main flow direction, resulting in a small
erosion rate.

@@

We next analyze the effect of erosion on the area fraction and the flow
rate.  In figure~\ref{fig:Eroding20flowrate}(a), we plot the area
fraction as a function of normalized time.  The general trend of the
area fraction resembles our previous work~\citep[see][figure
10(a)]{qua-moo2018}, but with a larger initial area fraction.  In
figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\citep[see][figure 10(b)]{qua-moo2018}, except that the initial
flow rate is an order of magnitude smaller because of the larger initial
area fraction.  Starting around normalized time $0.2$,
figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as an exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 4.32 \frac{t}{t_f} + c, 
\end{align}
where $t_f$ is the time required for all the bodies to erode, and $c$ is
a fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-4.32$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot, but with additional streamlines.
Since we use the Barycentric quadrature rule and a fourth-order
Runge-Kutta method, some streamlines are very close to the grains
without penetrating them.}
\end{center}
\end{figure}

In figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized  with the
streamlines.  In figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from figure~\ref{fig:Eroding20vort} and
plot 200 streamlines that are initialized at $x=-1$ and equally spaced
in $y \in (-1,1)$.  The streamlines are shown at five different times,
and the final plot is a zoom in of the lower right quadrant of the fifth
time step, but with additional streamlines.  Since we use a high-order
quadrature rule and time stepping method, we resolve streamlines that
come very close to the eroded bodies.  There are three clear regions
where the streamlines are most concentrated, corresponding to the
regions of highest velocity.  Two of these regions are located between
the bodies and the solid walls at $y=\pm 1$, and the third cuts through
the porous region with the upper part of the channel formed by bodies 1,
4, 6, and 8.  Since the flow is fastest in these regions, the shearing
is largest, and this causes the channels to continue to open fastest as
observed in figure~\ref{fig:Eroding20vort}.

Next, we use the $N_p = 1000$~\citep{bel-sal-rin1992} streamlines to
compute the tortuosity of the eroding geometry.  To compute the
tortuosity, we require the velocity at the inlet $x=-1$.  These
normalized velocities are plotted in figure~\ref{fig:Eroding20tort}(a)
for the eroded geometry at porosity $\phi = 62.9\%$
(figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of~\citet[see figure 4(a)]{mat-kha-koz2008}, except that our
cross-section, by construction, does not cut through any of the grains.
Next, in figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each streamline as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the
streamlines is 27\% longer than it would have been if the grains were
absent.  The average streamline is 9.79\% longer or equivalently the
tortuosity of the geometry is $1.098$.  Again, comparing the local
tortuosity to~\citet[see figure 4(b)]{mat-kha-koz2008}, the results are
qualitatively similar. However, since our initial cross-section does not
cut through the grains, the local tortuosity does not have any gaps.
Discontinuities in local tortuosity occur when nearby streamlines
diverge to circumvent a grain.  In figure~\ref{fig:Eroding20tort}(c), we
plot pairs of streamlines associated with the ten largest jumps in the
local tortuosity, with each pair of corresponding streamlines plotted in
the same color.

@@

In figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $T = 1.16$.  The tortuosity is computed with
both the length of the streamlines~\eqref{eqn:tortuosity1} (red stars)
and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and from applying
quadrature to compute the tortuosity.  As the bodies erode, wide
channels form where streamlines undergo only minor vertical deflections,
and this explains why the tortuosity eventually decreases with porosity.
We computed lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels} and found the power law that
minimizes the error.  The  black dashed line in
figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064},
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  Interestingly,
at the low porosities, the tortuosity initially increases. This increase
occurs because in the absence of erosion (left plot in
figure~\ref{fig:Eroding20vort}), many of the streamlines, such as those
initialized between bodies 15 \& 18, only perform minor deflections to
pass through the narrow regions, albeit, very slowly.  However, as
erosion starts to open the channels, the streamlines deflect into the
fast regions, such as the region above body 11, and this increases the
amount of vertical deflection, and therefore the tortuosity. While this
increase in tortuosity is interesting, in the next two examples we will
see that the tortuosity does not initially increase.

@@

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a) The tortuosity of an eroding
geometry initialized with 20 grains.  The tortuosity is calculated using
the Eulerian method~\eqref{eqn:tortuosity2} (blue dots) and Lagrangian
method~\eqref{eqn:tortuosity1} (red stars).  The red square corresponds
to the geometry in figure~\ref{fig:Eroding20tort}(c).  The dash line is
the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2064$, and
the corresponding root-mean-square error is $5.90 \times 10^{-3}$. (b)
The temporal evolution of $\sigma_\lambda$ as a function of time at
seven different porosities.  The dashed line, which has slope one,
corresponds to ballistic dispersion $\sigma_\lambda \sim t$ that occurs
if no grains are present. At early times, the particles undergo a
ballistic motion, but once they have traversed a few grains, the
spreading becomes super-dispersive with $\sigma_\lambda \sim t^\alpha$,
$\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of best fit
with slopes $\alpha = 1.06$ ($\phi=95.10\%$), $\alpha = 1.07$
($\phi=85.09\%$), $\alpha = 1.06$ ($\phi=75.15\%$), $\alpha = 0.97$
($\phi=65.09\%$), $\alpha = 0.78$ ($\phi=55.10\%$), $\alpha = 0.75$
($\phi=45.08\%$), and $\alpha = 0.56$ ($\phi=37.68\%$).  Values greater
than 1 are a result of using a least-squares fit for the tails of the
particle spreading.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and vorticity at four evenly spaced instances in time.
Initially, the smallest distance between pairs of bodies is $3.29 \times
10^{-4}$, and the smallest distance between the bodies and solid wall is
$4.50 \times 10^{-3}$.  At these distances, a resolution of
approximately $N_\iin = 27,000$ and $N_\out = 18,000$ discretization
points is required to satisfy the $5h$ threshold needed for the
trapezoid rule to achieve machine precision.  By applying the
Barycentric quadrature rule, we stably simulate erosion at the much
coarser resolution of $N_\iin = 256$ and $N_\out = 1024$ discretization
points.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results of dense packings of bodies eroding in
Stokes flow. Moreover, we analyze transport through the eroded
geometries by examining coarse-grained variables.  Each eroding body is
initially a circle with center $\cc_i$, $r_i$, and length $L_i = 2\upi
r_i$.  The center and radius are chosen at random, and the body is
accepted if it is contained in $[-1,1] \times [-1,1]$ and is
sufficiently separated from all other bodies.  Since the error of our
quadrature method is bounded independent of the geometry, we consider
bodies that are separated by less than 10\% of an arclength spacing.
The randomized method is repeated until the initial geometry reaches a
desired initial porosity.

For all simulations, we discretize each eroding grain with $N_\iin =
256$ points and the outer wall $\Gamma$ with $N_\out=1024$ points.  A
no-slip boundary condition is imposed on each eroding body $\gamma_i$,
and a Dirichlet boundary condition on $\Gamma$ is used to approximate a
far-field boundary condition. For all but the first example, the
Dirichlet boundary condition is a Hagen-Poiseuille flow, and the flow
rate is adjusted at each time step to maintain a constant pressure drop.
Since the fluid equations are linear, this is achieved by computing the
pressure near the inlet and outlet at each time step, and then scaling
the flow rate appropriately~\citep{qua-moo2018}.

The erosion rate loses regularity at stagnation points, which inevitably
leads to corner formation on the bodies.  As described in
section~\ref{sec:time} and our previous work~\citep{qua-moo2018}, we
ameliorate corner formation by introducing a curvature penalization term
with parameter $\eps$ and a Gaussian smoothing step with parameter
$\sigma$.  For all examples, we use the smoothing parameters
$\eps=15/256$ and $\sigma=10/256$, and the time step size is $\Delta t =
10^{-4}$.

In each of the examples, we plot the vorticity in the fluid bulk.  Since
the vorticity and the shear stress agree on the boundary of the eroding
grains, the vorticity is useful to visualize the erosion rate.  In
particular, the rate of erosion is largest in regions where the
magnitude of the vorticity is largest.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Consider a single eroding body close to a solid wall with the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. \citet{mit-spa2017} performed a similar
three-dimensional experiment using a second-order quadrature method.
Their initial body is a sphere with its center located $1.5$ radii above
the solid wall.  We initialize the two-dimensional eroding body with
radius $r = 0.4$, and we conduct numerical experiments where the initial
distance between the grain and the solid wall is $h$, $h/2$, and $h/10$,
where $h = 2\upi r/N_\iin$.  If we used the trapezoid rule and required
an error that is comparable to the Barycentric quadrature rule, the body
with an initial distance of $h/10$ from the solid wall would require
$6,400$ discretization points, and the outer wall would require $50,000$
discretization points.

In figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For all three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop.
However, in the top half of the body, there is a sudden increase in the
shear stress, and this leads to a region of high curvature.  This
behavior is also present in three dimensions~\citep[see figure
7(c)]{mit-spa2017}.  The biggest difference between the two- and
three-dimensional results is the presence of a recirculation zone.  In
three dimensions, there is no recirculation between the solid wall and
the spherical body~\citep{cha-feu2003}, but recirculation is possible in
two dimensions~\citep{chw-wu1975, hig1985}.  To visualize the flow, we
plot the vorticity of the final time step from figure~\ref{fig:NearWall}
in figure~\ref{fig:NearWall_vort}.  In these examples, a small
recirculation zone, both in size and magnitude, is present in the region
where the vorticity is smallest

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding grains with the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
imposed on $\Gamma$. The flow rate $U$ is chosen so that the average
pressure drop from $x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$
once all the grains have vanished.  The vorticity and grain
configuration at four equispaced times are shown in
figure~\ref{fig:Eroding20vort}.  Initially, several of the grains are
closer to the outer wall than the $5h$ threshold required for the
trapezoid rule to achieve machine precision.  In particular, the
distance between bodies 1, 6, 13, and 15 and the outer wall is $1.3h$,
$2.9h$, $2.8h$, and $1.3h$, respectively, where $h$ is the arclength
spacing of the outer wall $\Gamma$.  In addition, the distance between
several pairs of eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and
14 \& 18, is too small to be accurately resolved with the trapezoid
rule.  By using the Barycentric quadrature rule, the interaction between
these nearly-touching bodies is accurately resolved to the desired
accuracy, and erosion can be simulated until all the bodies have
vanished.

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.227\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow.  The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In the fourth frame,
bodies 4, 12, and 17 have completely vanished, and body 19 has almost
completely eroded.}
\end{center}
\end{figure}

Erosion causes the some of the pore sizes to quickly grow, and flat
faces develop along the regions of near contact.  This qualitative
behavior is seen in figure~\ref{fig:Eroding20vort} between bodies 3 \&
4, 15 \& 16, and was also observed in previous work~\citep{qua-moo2018}.
However, by resolving the interaction between bodies that are much
closer together, we observe that very little erosion occurs between
certain pairs of bodies, at least initially.  For instance the opening
between bodies 1 \& 6, 3 \& 9, and 5 \& 13 grow much slower than the
opening between bodies 15 \& 16.  A common feature of the pores that
grow slowly is that they are nearly perpendicular to the main flow
direction, resulting in a small erosion rate.

@@

We next analyze the effect of erosion on the area fraction and the flow
rate.  In figure~\ref{fig:Eroding20flowrate}(a), we plot the area
fraction as a function of normalized time.  The general trend of the
area fraction resembles our previous work~\citep[see][figure
10(a)]{qua-moo2018}, but with a larger initial area fraction.  In
figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\citep[see][figure 10(b)]{qua-moo2018}, except that the initial
flow rate is an order of magnitude smaller because of the larger initial
area fraction.  Starting around normalized time $0.2$,
figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as an exponential law.  The line of
best fit is
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from figure~\ref{fig:Eroding20vort}. The streamlines are
initially equispaced at $(-1,y)$, where $y \in (-1,1)$. The first five
snapshots are evenly spaced in time.  The bottom right frame is a zoom
in of the fifth snapshot, but with additional streamlines.  Since we use
the Barycentric quadrature rule and a fourth-order Runge-Kutta method,
streamlines that are very close to the grains are resolved.}
\end{center}
\end{figure}

In figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to the outlet where the velocity and
vorticity, and therefore erosion rate, are much larger relative to other
regions.  These channels can be further visualized  with the
streamlines.  In figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from figure~\ref{fig:Eroding20vort} and
plot 200 streamlines that are initially equispaced along $(-1,y)$, where
$y \in (-1,1)$.  The streamlines are shown at five different times, and
the final plot is a zoom in of the lower right quadrant of the fifth
time step, but with additional streamlines.  Since we use a high-order
quadrature rule and time stepping method, we resolve streamlines that
come very close to the eroded bodies.  There are three clear regions
where the streamlines are most concentrated, corresponding to the
regions of highest velocity.  Two of these regions are located between
the bodies and the solid walls at $y=\pm 1$, and the third cuts through
the porous region with the upper part of the channel formed by bodies 1,
4, 6, and 8.  Since the flow is fastest in these regions, the shearing
is largest, and this causes the channels to continue to open fastest as
observed in figure~\ref{fig:Eroding20vort}.

Next, we use the $N_p = 1000$~\citep{bel-sal-rin1992} streamlines to
compute the tortuosity of the eroding geometry.  To compute the
tortuosity, we require the velocity at the inlet $x=-1$.  These
normalized velocities are plotted in figure~\ref{fig:Eroding20tort}(a)
for the eroded geometry at porosity $\phi = 62.9\%$
(figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar those
of~\citet[see figure 4(a)]{mat-kha-koz2008}, except that our
cross-section, by construction, does not cut through any of the grains.
Next, in figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each streamline as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the
streamlines is 27\% longer than it would have been if the grains were
absent.  The average streamline is 9.79\% longer or equivalently the
tortuosity of the geometry is $1.098$.  Again, comparing the local
tortuosity to~\citet[see figure 4(b)]{mat-kha-koz2008}, the results are
qualitatively similar. However, since our initial cross-section does not
cut through the grains, the local tortuosity does not have any gaps.
Discontinuities in local tortuosity occur when nearby streamlines
diverge to circumvent a grain.  In figure~\ref{fig:Eroding20tort}(c), we
plot pairs of streamlines associated with the ten largest jumps in the
local tortuosity, with each pair of corresponding streamlines plotted in
the same color.

@@

In figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $T = 1.16$.  The tortuosity is computed with
both the length of the streamlines~\eqref{eqn:tortuosity1} (red stars)
and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and from applying
quadrature to compute the tortuosity.  As the bodies erode, wide
channels form where streamlines undergo only minor vertical deflections,
and this explains why the tortuosity eventually decreases with porosity.
We computed lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels} and found that the power law
minimizes the error.  The  black dashed line in
figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064},
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  Interestingly,
at the low porosities, the tortuosity initially increases. This increase
occurs because in the absence of erosion (left plot in
figure~\ref{fig:Eroding20vort}), many of the streamlines, such as those
initialized between bodies 15 \& 18, only perform minor deflections to
pass through the narrow regions, albeit, very slowly.  However, as
erosion starts to open the channels, the streamlines deflect into the
fast regions, such as the region above body 11, and this increases the
amount of vertical deflection, and therefore the tortuosity. While this
increase in tortuosity is interesting, in the next two examples we will
see that the tortuosity does not initially increase.

@@

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a) The tortuosity of an eroding
geometry initialized with 20 grains.  The tortuosity is calculated using
the Eulerian method~\eqref{eqn:tortuosity2} (blue dots) and Lagrangian
method~\eqref{eqn:tortuosity1} (red stars).  The red square corresponds
to the geometry in figure~\ref{fig:Eroding20tort}(c).  The dash line is
the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2064$, and
the corresponding root-mean-square error is $5.90 \times 10^{-3}$. (b)
The temporal evolution of $\sigma_\lambda$ as a function of time at
seven different porosities.  The dashed line, which has slope one,
corresponds to ballistic dispersion $\sigma_\lambda \sim t$ that occurs
if no grains are present. At early times, the particles undergo a
ballistic motion, but once they have traversed a few grains, the
spreading becomes super-dispersive with $\sigma_\lambda \sim t^\alpha$,
$\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of best fit
with slopes $\alpha = 1.06$ ($\phi=95.10\%$), $\alpha = 1.07$
($\phi=85.09\%$), $\alpha = 1.06$ ($\phi=75.15\%$), $\alpha = 0.97$
($\phi=65.09\%$), $\alpha = 0.78$ ($\phi=55.10\%$), $\alpha = 0.75$
($\phi=45.08\%$), and $\alpha = 0.56$ ($\phi=37.68\%$).  Values greater
than 1 result from using a least-squares fit for the tails of the
particle spreading.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{Bryan is here}
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and vorticity at four evenly spaced instances in time.
Initially, the smallest distance between pairs of bodies is $3.29 \times
10^{-4}$, and the smallest distance between the bodies and solid wall is
$4.50 \times 10^{-3}$.  At these distances, a resolution of
approximately $N_\iin = 27,000$ and $N_\out = 18,000$ discretization
points is required to satisfy the $5h$ threshold needed for the
trapezoid rule to achieve machine precision.  By applying the
Barycentric quadrature rule, we stably simulate erosion at the much
coarser resolution of $N_\iin = 256$ and $N_\out = 1024$ discretization
points.

\begin{figure}
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In addition to the
channels that develop between the bodies and the solid walls, erosion
leads to two main channels through the geometry---one in the top half
and one near the middle.}
\end{figure}
